Selecting and ranking interesting pairs of cyclic temporal granularities

### 2020-12-10 paper/main.Rmd
tuning paramter section update

bibliography: bibliography.bib
preamble: >
  \usepackage{mathtools,amssymb,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \usepackage{algorithm}
  \usepackage{algorithmicx}
  \usepackage{algpseudocode}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE,
        tinytex.verbose = TRUE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(viridis)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)

@@

# Median Maximum Pairwise Distances (MMPD)

@@

## Computation

@@

### Methodology

@@

```{r distance-explain, fig.pos="p", fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets. The dotted lines represent the distances between different categories. Panel 1) and 2) show the between-facet distances. Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively. When categories are ordered, distances should only be considered for consecutive x-axis categories. Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}
knitr::include_graphics(here::here("paper/Figs/dist_explain.png"))
```

From Section \ref{sec:idea}, the idea is to put more weightage on within-facet distances than between-facet distances. Hence, for a suitable tuning parameter $\lambda > 1$, the pairwise distances $d_{(jk, j'k')}$ are transformed based on the distance type as follows:

@@

<!-- The maximum weighted pairwise distances are defined as: -->
<!-- $$MEPD = max_{j, j', k, k'}(d*_{(j,k), (j', k')}) -->
<!-- \forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$ -->
<!-- where,  -->

@@

>>>>

bibliography: bibliography.bib
preamble: >
  \usepackage{mathtools,booktabs,amsthm,todonotes,colortbl}
  \def\mod{~\text{mod}~}
  \newtheorem{definition}{Definition}
  \usepackage{mathptmx}
  \usepackage{caption}
  \usepackage{algorithm}
  \usepackage{algorithmicx}
  \usepackage{algpseudocode}
  \DeclareCaptionStyle{italic}{labelfont={bf},textfont={it},labelsep=colon}
  \captionsetup[figure]{style=italic,format=hang,singlelinecheck=true}
  \captionsetup[table]{style=italic,format=hang,singlelinecheck=true}
  \def\novspacing{\setlength{\aboverulesep}{0pt}\setlength{\belowrulesep}{0pt}}
  \def\vspacing{\setlength{\aboverulesep}{0.4ex}\setlength{\belowrulesep}{0.65ex}}
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
---

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE,
        tinytex.verbose = FALSE)
library(knitr)
library(tidyverse)
library(lubridate)
library(lvplot)
library(ggridges)
library(viridis)
library(tsibble)
library(gravitas)
library(ggpubr)
library(readr)
library(kableExtra)
library(distributional)
library(ggplot2)
library(sugrrants)
library(here)

@@

# Distance measure

@@

@@

### Definitions

@@

```{r distance-explain, fig.cap = "Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets. The dotted lines represent the distances between different categories. Panel 1) and 2) show the between-facet distances. Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively. When categories are ordered, distances should only be considered for consecutive x-axis categories. Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}."}
knitr::include_graphics(here::here("paper/Figs/dist_explain.png"))
```

From Section \ref{sec:idea}, the idea is to put more weights on within-facet distances than between-facet distances. Hence, for a suitable tuning parameter $\lambda > 1$, the pairwise distances $d_{(jk, j'k')}$ are transformed based on the distance type as follows:

@@

The maximum weighted pairwise distances are defined as:
$$WPD = max_{j, j', k, k'}(d*_{(j,k), (j', k')})
\forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}$$.

@@

D1 - Design null: 
D2 - Design vary_f:
D3 - Design var_x:
B4 - Design vary_all:

## Choice of parameters of the distribution

Under null distribution

Null design only where all combinations are equal for fixed values of $\mu$ and $\sigma$
Dist 1: $N(0,1)$ individual graph
Dist 2: $N(\mu, 1)$ (faceted with different values of $\mu$: not yet just for one value now)
Dist 3: $N(0, \sigma)$ (faceted with different values of $\sigma$: not yet just for one value now)
Dist 4: $N(\mu, \sigma)$ faceted with different values of mu and sigma, for all levels of x and facet to see how value of wpd changes with that of change in parameters.

### Choice of alternate designs

A normal distribution is taken for this purpose. We would calibrate the values of $WPD$ for different designs illustrated in Section \ref{sec:idea}.

$$\mu_{j.} = \mu + j\omega$$
or, $$\mu_{.k} = \mu + k\omega$$ where,   
$j$ and $k$ are the index of the x and facet category respectively 
$\mu$ is the mean of the first category and $\omega$ is the increment from the starting value.

mean = 0
range_mean = c(0, 1, 3,  5, 7, 15, 20)
sd = 1
range_sd = c(1, 3,  5, 7, 15, 20)
range_w =  c(1, 2, 3, 4, 5)
range_lambda = seq(0.1, 0.9, 0.05)

## Parameters of the distribution

## Choice of tuning parameters

Objective: Choice of optimal tuning parameter $lambda$ with the help of two alternate designs (vary_x and vary_f)  for different nx and nfacets. The lambda for which the two designs intersect have to be chosen to put appropriate weights for within-facet and between-facet distances.

Assumptions: The initial combination has a $N(\mu,\sigma)$ distribution and then every combination increases as follows:
- $$\mu_{j.} = \mu + j\omega$$ (for design vary_x)
- $$\mu_{.k} = \mu + k\omega$$ (for design vary_f)
Value of the parameters chosen for this simulation are $\mu = 0$, $\sigma = 1$ and $\omega = {1, 2, \dots, 10}$

Fixing omega, where the design interacts

```{r tuning-one-omega}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

How value of lambda changes with increasing omega

```{r tuning-parameter-intersection}
knitr::include_graphics(here::here("simulations/tuning_param/figs/intersection_plot.png"))
```

How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation?

```{r lambda-omega-mean}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_mean_3_2.png"))
```

How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean?

```{r lambda-omega-sd}
knitr::include_graphics(here::here("simulations/result_report/lambda_omega_sd_3_2.png"))
```

### 2020-12-10 paper/main.tex
tuning paramter section update

\hypertarget{median-maximum-pairwise-distances-mmpd}{%
\section{Median Maximum Pairwise Distances (MMPD)}\label{median-maximum-pairwise-distances-mmpd}}

@@

\hypertarget{idea}{%
\subsection{Idea}\label{idea}}

@@

\hypertarget{computation}{%
\subsection{Computation}\label{computation}}

@@

\hypertarget{algorithm}{%
\subsubsection{Algorithm}\label{algorithm}}

Consider two cyclic granularities \(C_i\) and \(C_j\), such that \(C_i\) maps index set to a set \(\{A_k \mid k=1,\dots,K\}\) and \(C_j\) maps index set to a set \(\{B_\ell \mid \ell =1,\dots,L\}\). Here, \(A_k\) and \(B_\ell\) are the levels/categories corresponding to \(C_i\) and \(C_j\) respectively. Let \(S_{k\ell}\) be a subset of the index set such that for all \(s \in S_{k\ell}\), \(C_i(s) = A_k\) and \(C_j(s) = B_\ell\). There are \(KL\) such data subsets, one for each combination of levels (\(A_k\), \(B_\ell\)). Moreover, consider that in the graphical space, \(C_i\) is mapped to facets and \(C_j\) is mapped to x-axis.

The algorithm employed for computing the distance measure is summarized as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Fix harmony pair \((C_i, C_j)\).
\item
  Fix \(k\). Then there are \(L\) groups corresponding to level \(A_k\) of \(C_i\).
\item
  Compute \(m = \binom{L}{2}\) pairwise distances between distributions of \(L\) unordered levels and \(m = L-1\) pairwise distances for \(L\) ordered categories.
\item
  Identify maximum within the \(m\) computed distances.
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{4}
\item
  Use Steps 1-4 to compute maximum distance for \(\forall k \in \{1, 2, \ldots, K\}\).
\item
  Compute the distance measure MMPD\_raw = median \((M_1, M_2, \dots, M_K)\).
\end{enumerate}

\begin{algorithm}[!thb]
    \caption{Calculation for a raw distance measure between two cyclic granularities $A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$ with $A$ placed across x-axis and $B$ across facets.}\label{alg:scoreopt}.
    \begin{algorithmic}[1]
        \Procedure{RawMMPD}{$A = \{ a_j: j = 1, 2, \dots, J\}$, $B = \{ b_k: k = 1, 2, \dots, K\}$, $v = \{ v_t: t = 1, 2, \dots, T\}$}.
        \For{$k=1:K$, $j=1:J$}
        \State Find distances between pairs of all possible combinations of categories $(a_jb_k,a_j'b_k')$ by computing JSD between quantiles of the measured variable $q(v)$ across these combinations.
        \State $d \gets JSD(\tilde{q(v)_{a_{j}b_{k}}},\tilde{q(v)_{a_{j}'b_{k}'}})$
        \If {$b_k = b_k'$}
        \State $d* \gets \lambda d$  \Comment{upweight within-facet distances}
        \Else
        \State $d* \gets 1/\lambda d $ \Comment{downweight across-facet distances}
        \EndFor
        \State Set the raw distance measure as $max(d*)$ where max is taken over all $j, j', k, k'$.

        \EndProcedure
    \end{algorithmic}
\end{algorithm}

A stronger measure ``max'' is chosen for aggregating x-axis categories compared to ``median'' for aggregating facet categories as the measure is intended to put more importance in pointing towards distributional differences between x-axis categories than for facet categories.

@@

@@

@@

\hypertarget{algorithm-1}{%
\subsection{Algorithm}\label{algorithm-1}}

>>>>

\hypertarget{distance-measure}{%
\section{Distance measure}\label{distance-measure}}

@@

\hypertarget{sec:idea}{%
\subsection{Idea}\label{sec:idea}}

@@

@@

\hypertarget{definitions}{%
\subsubsection{Definitions}\label{definitions}}

Consider two cyclic granularities \(A\) and \(B\), such that \(A = \{ a_j: j = 1, 2, \dots, J\}\) and \(B = \{ b_k: k = 1, 2, \dots, K\}\) with \(A\) placed across x-axis and \(B\) across facets. Let the pairwise distances between pairs \((a_{j} b_{k}, a_{j'}, b_{k})\) be denoted as \(d_{(jk, j', k')} = JSD(a_{j}b_{k}, a_{j'}b_{k'})\). Pairwise distances could be within-facets or between-facets. Figure \ref{fig:distance-explain} illustrates how the within-facet or between-facet distances are defined. Pairwise distances are within-facets (\(d_{w}\)) when \(b_{k} = b_{k'}\), that is, between pairs of the form \((a_{j}b_{k}, a_{j'}b_{k})\) as shown in panel (3) of Figure \ref{fig:distance-explain}. If categories are ordered (like all temporal cyclic granularities), then only distances between pairs where \(a_{j'} = (a_{j+1})\) are considered (panel (4)). Pairwise distances are between-facets (\(d_{b}\)) when they are considered between pairs of the form \((a_{j}b_{k}, a_{j}b_{k'})\).

{\centering \includegraphics[width=\textwidth]{/Users/sgup0008/Documents/paper-hakear/paper/Figs/dist_explain}

\caption{Within and between-facet distances shown for two cyclic granularities A and B, where A is mapped to x-axis and B is mapped to facets. The dotted lines represent the distances between different categories. Panel 1) and 2) show the between-facet distances. Panel 3) and 4) are used to illustrate within-facet distances when categories are un-ordered or ordered respectively. When categories are ordered, distances should only be considered for consecutive x-axis categories. Between-facet distances are distances between different facet levels for the same x-axis category, for example, distances between {($a_1$,$b_1$) and ($a_1$, $b_2$)} or {($a_1$,$b_1$) and ($a_1$, $b_3$)}.}\label{fig:distance-explain}
\end{figure}

From Section \ref{sec:idea}, the idea is to put more weights on within-facet distances than between-facet distances. Hence, for a suitable tuning parameter \(\lambda > 1\), the pairwise distances \(d_{(jk, j'k')}\) are transformed based on the distance type as follows:

\begin{equation}
d*_{(j,k), (j'k')} =
\begin{cases}
\lambda d_{(j,k), (j'k')},& \text{if } d = d_w\\
(1-\lambda)  d_{(j,k), (j', k')},              & \text{if } d = d_b\\
\end{cases}
\end{equation}

The maximum weighted pairwise distances are defined as:
\[WPD = max_{j, j', k, k'}(d*_{(j,k), (j', k')})
\forall j, j' \in \{1, 2, \dots, J\}, k, k' \in \{1, 2, \dots, K\}\].

--\textgreater{}

@@

D1 - Design null:
D2 - Design vary\_f:
D3 - Design var\_x:
B4 - Design vary\_all:

\hypertarget{choice-of-parameters-of-the-distribution}{%
\subsection{Choice of parameters of the distribution}\label{choice-of-parameters-of-the-distribution}}

Under null distribution

Null design only where all combinations are equal for fixed values of \(\mu\) and \(\sigma\)
Dist 1: \(N(0,1)\) individual graph
Dist 2: \(N(\mu, 1)\) (faceted with different values of \(\mu\): not yet just for one value now)
Dist 3: \(N(0, \sigma)\) (faceted with different values of \(\sigma\): not yet just for one value now)
Dist 4: \(N(\mu, \sigma)\) faceted with different values of mu and sigma, for all levels of x and facet to see how value of wpd changes with that of change in parameters.

\hypertarget{choice-of-alternate-designs}{%
\subsubsection{Choice of alternate designs}\label{choice-of-alternate-designs}}

A normal distribution is taken for this purpose. We would calibrate the values of \(WPD\) for different designs illustrated in Section \ref{sec:idea}.

\[\mu_{j.} = \mu + j\omega\]
or, \[\mu_{.k} = \mu + k\omega\] where,\\
\(j\) and \(k\) are the index of the x and facet category respectively
\(\mu\) is the mean of the first category and \(\omega\) is the increment from the starting value.

mean = 0
range\_mean = c(0, 1, 3, 5, 7, 15, 20)
sd = 1
range\_sd = c(1, 3, 5, 7, 15, 20)
range\_w = c(1, 2, 3, 4, 5)
range\_lambda = seq(0.1, 0.9, 0.05)

\hypertarget{parameters-of-the-distribution}{%
\subsection{Parameters of the distribution}\label{parameters-of-the-distribution}}

\hypertarget{choice-of-tuning-parameters}{%
\subsection{Choice of tuning parameters}\label{choice-of-tuning-parameters}}

Objective: Choice of optimal tuning parameter \(lambda\) with the help of two alternate designs (vary\_x and vary\_f) for different nx and nfacets. The lambda for which the two designs intersect have to be chosen to put appropriate weights for within-facet and between-facet distances.

Assumptions: The initial combination has a \(N(\mu,\sigma)\) distribution and then every combination increases as follows:
- \[\mu_{j.} = \mu + j\omega\] (for design vary\_x)
- \[\mu_{.k} = \mu + k\omega\] (for design vary\_f)
Value of the parameters chosen for this simulation are \(\mu = 0\), \(\sigma = 1\) and \(\omega = {1, 2, \dots, 10}\)

Fixing omega, where the design interacts

\begin{center}\includegraphics[width=\textwidth]{/Users/sgup0008/Documents/paper-hakear/simulations/tuning_param/figs/intersection_plot} \end{center}

How value of lambda changes with increasing omega

\begin{center}\includegraphics[width=\textwidth]{/Users/sgup0008/Documents/paper-hakear/simulations/tuning_param/figs/intersection_plot} \end{center}

How value of lambda changes with increasing omega and mean for fixed nx, nfacet and standard deviation?

\begin{center}\includegraphics[width=\textwidth]{/Users/sgup0008/Documents/paper-hakear/simulations/result_report/lambda_omega_mean_3_2} \end{center}

How value of lambda changes with increasing omega and sd for fixed nx, nfacet and mean?

\begin{center}\includegraphics[width=\textwidth]{/Users/sgup0008/Documents/paper-hakear/simulations/result_report/lambda_omega_sd_3_2} \end{center}

@@

\hypertarget{methodology-1}{%
\subsection{Methodology}\label{methodology-1}}

@@

\hypertarget{algorithm}{%
\subsection{Algorithm}\label{algorithm}}
